Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-22 latex/article.tex
sect4

\paragraph{Hierarchical decomposition}  As discussed above,

\paragraph{Dynamic refinement}

\begin{itemize}
\item   Variante 1 : hierarchique avec précalcul cells
\item   Variante 2 : reg avec autoraffinement
\end{itemize}

@@

\appendix
\section{Hierarchical Probing Algorithm}

A {\em cell} is a tuple $(k,x,y,z)$, which represents a region of the space of size $2^k \times
2^k \times 2^k$, characterized by its integer coordinates $x,y,z$,
with $0 \le x < 2^k$, $0 \le y < 2^k$, $0 \le z < 2^k$. Cells forms an
octree decomposition of a cubic space. We use functions Up, Down and
Next to navigate between cells.

\begin{algorithm}
\KwIn{Integers $p,q,r$ \tcp*{the moment orders, with $0 \le p+q+r \le 2$}}
\KwIn{Integer $k$ \tcp*{$(2^k)^3$ is the size of the digital shape image}}
\KwIn{Mipmap $V$ \tcp*{array of $k+1$ images of sizes $(2^k)^3,  (2^{k-1})^3, \ldots,  1^3$}}
\KwIn{Integers $x_0,y_0,z_0$, Real $r$ \tcp*{Ball radius $r$ and center $(x_0,y_0,z_0)$}}
\KwOut{Real $m$ \tcp*{estimation of the $p,q,r$-moment of $X \cap B_r(x_0,y_0,z_0)$}}
\KwData{Cell $c :=  (k,0,0,0)$ \tcp*{Starts from biggest cell}}
\KwData{Integer $n := 2^k$ \tcp*{Size of each cell}}
\KwData{Real $d,\delta,l$ \tcp*{variables for intermediate computations}}
\Begin{
    m := 0 \;
    \Repeat{$c[0] = k$}{
      \tcp{Distance between cell and ball centers}
      $d := \frac{1}{2}\| 2^k(2c[1]+1,2c[2]+1,2c[3]+1) - (2x_0+1,2y_0+1,2z_0+1) \|_2$\;
      $\delta := \frac{\sqrt{3}}{2}2^{c[0]}$ \tcp*{half-length of cell diagonal}
      \If(\tcp*[f]{Is it a unit cell ?}){$c[0] = 0$}{
        \If(\tcp*[f]{Is cell center inside ball ?}){$d^2 \le r^2$}{
          $m := m + V[c] * (2^k c[1])^p * (2^k c[2])^q * (2^k c[3])^r$ \;
        }
        $c := \textsc{Next}(c)$ \tcp*{Go to next cell}
      } \Else{
        $l := \max(r-\delta,0)$\;
        \If(\tcp*[f]{Is cell completely inside ball ?}){$d^2 < l^2$}{
          $m := m + V[c] * (2^k c[1])^p * (2^k c[2])^q * (2^k c[3])^r$\;
          $c := \textsc{Next}(c)$ \tcp*{Go to next cell}
        }\ElseIf(\tcp*[f]{Is cell outside ball ?}){$d^2 > (r+\delta)^2$}{
          $c := \textsc{Next}(c)$ \tcp*{Go to next cell}
        }\lElse(\tcp*[f]{Go to a finer cell}){$c := \textsc{Down}(c)$}
      }
    }
    \Return{m}\;
  }
  \caption{Derecursified hierarchical algorithm for computing the
    $p,q,r$-moment of set $X \cap B_r(x_0,y_0,z_0)$, given a mipmap
    $V$ that represents the volume of a shape $X$ in each cell, and
    ball parameters $x_0,y_0,z_0,r$.}
\end{algorithm}

\begin{function}
  \caption{Up( Cell $c$ ) : Cell}
  \Return{Cell( $c[0]+1$, $c[1]/2$, $c[2]/2$, $c[3]/2$ )}
\end{function}
\begin{function}
  \caption{Down( Cell $c$ ) : Cell}
  \Return{Cell( $c[0]-1$, $c[1]*2$, $c[2]*2$, $c[3]*2$ )}
\end{function}
\begin{function}
  \caption{Next( Cell $c$ ) : Cell}
  \lWhile{$\mathrm{Odd}(c[1])$ and $\mathrm{Odd}(c[2])$ and $\mathrm{Odd}(c[3])$}{$c := \mathrm{Up}(c)$}
  \lIf{$\mathrm{Even}(c[1])$}{$c[1] := c[1]+1$}
  \Else{$c[1] := c[1] - 1$\;
    \lIf{$\mathrm{Even}(c[2])$}{$c[2] := c[2]+1$}
    \Else{$c[2] := c[2] - 1$\;
      $c[3] := c[3]+1$}
  }
  \Return{c}
\end{function}

>>>>

\paragraph{Hierarchical Decomposition.}  As discussed above, best
results in terms of texture fetch can be obtained by the approach
presented in Fig.~\ref{fig:approx2}-$b$. At each vertex, the
computation algorithm would consist in a recursive decomposition of
the domain such that a cell is subdivided if it does not belong to
$B_R(\vx)\cap\Shape$. However, on GPU hardware, this approach is
highly inefficient since the hardware optimizes the parallelism only
when the micro-program described in the shader has a predictive
execution flow. Hence, implementing the recursive algorithm (or its
de-recursified version) requires a lot of branches (conditional
\emph{if} tests) in the flow. Instead of implementing this approach,
we have considered a fast hierarchical approximation following
Fig.~\ref{fig:approx2}-$a$: for a given radius $R$, we precompute a
hierachical decomposition of $B_R$. Such decomposition is made of
octree cells at different resolutions. At a given MC vertex $\vx$, the
algorithm becomes simple since we just scan  the ball cells and estimate
the area (or other moments) from the mipmap values associated to this
cell. Note that cells in the $B_R$ decomposition may not be aligned
wuth mipmap cells. However, we can use GPU mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells on
$B_r$ is in $O(\log{R})$.

\todo[inline]{phrase sur approx si on met}

\paragraph{Dynamic Refinement.} In this approach, we want to optimize
the interactivity and the execution flow or parallelism on GPU. The
integration is simply computed using a regular grid at different
resolution $l$ (Fig. \ref{fig:approx}-$b$ and $c$). The shader code
becomes trivial (simple triple loop) and a lot of texture fetched are
involved, but interactivity can be easily obtained. Indeed, we
consider the following multi-pass approach:
\begin{enumerate}
\item When the surface geometry has been computed,we set the current
  level $l$ at an high level $l\EqDef l_{max}$ of the mipmap
  pyramid.
\item We compute the integrals (and the curvature tensor) at the
  mipmap level $l$ and  send the estimated quantities to the
  visualization stage.
\item If the user changes the camera position or the computation
  parameters, we return to step 1.
\item If the current framerate is above a given threshold, we decrease
  $l$ and return to step 2 if the final $l_0$ level has not been
  reached yet.
\end{enumerate}
A consequence of the multi-pass approach is that is there is no
interaction (camera settings, parameters), the GPU automatically
refine the estimation. Even if step 2 is quite expensive, in
$O\left(\left(\frac{R}{2^l}\right)^3\right)$, the interactive control
 in step 3 considerably improves the interactive exploration of the
 tensor with fast preliminary approximations which are quickly refined.

@@

%% \appendix
%% \section{Hierarchical Probing Algorithm}

%% A {\em cell} is a tuple $(k,x,y,z)$, which represents a region of the space of size $2^k \times
%% 2^k \times 2^k$, characterized by its integer coordinates $x,y,z$,
%% with $0 \le x < 2^k$, $0 \le y < 2^k$, $0 \le z < 2^k$. Cells forms an
%% octree decomposition of a cubic space. We use functions Up, Down and
%% Next to navigate between cells.

%% \begin{algorithm}
%% \KwIn{Integers $p,q,r$ \tcp*{the moment orders, with $0 \le p+q+r \le 2$}}
%% \KwIn{Integer $k$ \tcp*{$(2^k)^3$ is the size of the digital shape image}}
%% \KwIn{Mipmap $V$ \tcp*{array of $k+1$ images of sizes $(2^k)^3,  (2^{k-1})^3, \ldots,  1^3$}}
%% \KwIn{Integers $x_0,y_0,z_0$, Real $r$ \tcp*{Ball radius $r$ and center $(x_0,y_0,z_0)$}}
%% \KwOut{Real $m$ \tcp*{estimation of the $p,q,r$-moment of $X \cap B_r(x_0,y_0,z_0)$}}
%% \KwData{Cell $c :=  (k,0,0,0)$ \tcp*{Starts from biggest cell}}
%% \KwData{Integer $n := 2^k$ \tcp*{Size of each cell}}
%% \KwData{Real $d,\delta,l$ \tcp*{variables for intermediate computations}}
%% \Begin{
%%     m := 0 \;
%%     \Repeat{$c[0] = k$}{
%%       \tcp{Distance between cell and ball centers}
%%       $d := \frac{1}{2}\| 2^k(2c[1]+1,2c[2]+1,2c[3]+1) - (2x_0+1,2y_0+1,2z_0+1) \|_2$\;
%%       $\delta := \frac{\sqrt{3}}{2}2^{c[0]}$ \tcp*{half-length of cell diagonal}
%%       \If(\tcp*[f]{Is it a unit cell ?}){$c[0] = 0$}{
%%         \If(\tcp*[f]{Is cell center inside ball ?}){$d^2 \le r^2$}{
%%           $m := m + V[c] * (2^k c[1])^p * (2^k c[2])^q * (2^k c[3])^r$ \;
%%         }
%%         $c := \textsc{Next}(c)$ \tcp*{Go to next cell}
%%       } \Else{
%%         $l := \max(r-\delta,0)$\;
%%         \If(\tcp*[f]{Is cell completely inside ball ?}){$d^2 < l^2$}{
%%           $m := m + V[c] * (2^k c[1])^p * (2^k c[2])^q * (2^k c[3])^r$\;
%%           $c := \textsc{Next}(c)$ \tcp*{Go to next cell}
%%         }\ElseIf(\tcp*[f]{Is cell outside ball ?}){$d^2 > (r+\delta)^2$}{
%%           $c := \textsc{Next}(c)$ \tcp*{Go to next cell}
%%         }\lElse(\tcp*[f]{Go to a finer cell}){$c := \textsc{Down}(c)$}
%%       }
%%     }
%%     \Return{m}\;
%%   }
%%   \caption{Derecursified hierarchical algorithm for computing the
%%     $p,q,r$-moment of set $X \cap B_r(x_0,y_0,z_0)$, given a mipmap
%%     $V$ that represents the volume of a shape $X$ in each cell, and
%%     ball parameters $x_0,y_0,z_0,r$.}
%% \end{algorithm}

%% \begin{function}
%%   \caption{Up( Cell $c$ ) : Cell}
%%   \Return{Cell( $c[0]+1$, $c[1]/2$, $c[2]/2$, $c[3]/2$ )}
%% \end{function}
%% \begin{function}
%%   \caption{Down( Cell $c$ ) : Cell}
%%   \Return{Cell( $c[0]-1$, $c[1]*2$, $c[2]*2$, $c[3]*2$ )}
%% \end{function}
%% \begin{function}
%%   \caption{Next( Cell $c$ ) : Cell}
%%   \lWhile{$\mathrm{Odd}(c[1])$ and $\mathrm{Odd}(c[2])$ and $\mathrm{Odd}(c[3])$}{$c := \mathrm{Up}(c)$}
%%   \lIf{$\mathrm{Even}(c[1])$}{$c[1] := c[1]+1$}
%%   \Else{$c[1] := c[1] - 1$\;
%%     \lIf{$\mathrm{Even}(c[2])$}{$c[2] := c[2]+1$}
%%     \Else{$c[2] := c[2] - 1$\;
%%       $c[3] := c[3]+1$}
%%   }
%%   \Return{c}
%% \end{function}
